### 2024-02-16 code/process.Rmd
work on analysis

---
title: "251 Rescues"
output:
  html_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(viridis)
library(Replicate)
library(metafor)
library(esc)
library(here)
library(brms)
library(rstan)
library(googledrive)
library(glmnet)
library(tidybayes)
library(ggstance)
library("lattice")
library(reshape2)
library(ggrepel)
library(ggthemes)
library(knitr)
library(cowplot)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

theme_set(theme_bw())

model_location="code/models"
```

# Pull data

Download

```{r, eval=F}
f <- googledrive::as_dribble("https://docs.google.com/spreadsheets/d/12A4DblSbX_0tHP1mTVJhjNboI1YDngidkNWiSy0sN0g/edit#gid=0")
googledrive::drive_download(f, path=here("data","raw_data.xlsx"), overwrite=T)

raw_expts <- readxl::read_xlsx(here("data","raw_data.xlsx"), sheet="expt-level", skip=0)

raw_orig <- readxl::read_xlsx(here("data", "raw_data.xlsx"), sheet="original")

ready <- raw_expts |> left_join(raw_orig) |> write_csv(here("data", "combined_data.csv"))

```

Read in cache.

```{r}
d <- read_csv(here("data", "combined_data.csv")) |> 
  select(target_lastauthor_year, type, on_turk, repeated_measure, N, raw_stat, 
         same_direction, replication_score, closeness, subfield, target_year, stanford_internal, 
         open_data, open_materials, within_between, single_vignette)

```

# Parsing

We parse out values from the raw stats.

```{r}
source(here("code","helper","parse_stats.R"))
```

```{r}
parsed_d <- d |> 
  mutate(raw_stat=gsub(" ","",raw_stat),
         calc=pmap(list(raw_stat, within_between,N), do_parsing)) |> 
  unnest(cols=c(calc), names_sep="_") |> 
    mutate(
    calc_d_calc=case_when(
      type=="original" ~ abs(calc_d_calc),
      same_direction=="yes" ~ abs(calc_d_calc),
      same_direction=="no" ~ -abs(calc_d_calc),
      T ~ as.numeric(NA)
      ),
    calc_ES=case_when(
      type=="original" ~ abs(calc_ES),
      same_direction=="yes" ~ abs(calc_ES),
      same_direction=="no" ~ -abs(calc_ES),
      T ~ as.numeric(NA)
      ),
    type=factor(type, levels=c("original", "rep1", "rescue", "additional"))
    ) |> 
  rowwise()
```

## what didn't parse

Check that nothing that has a stat input and doesn't get an ES out.

```{r}

parsed_d |> filter(is.na(calc_ES)&(is.na(calc_d_calc))) |>  select(target_lastauthor_year, type, raw_stat)

parsed_d |> filter(is.na(calc_SE)&(is.na(calc_d_calc_se))) |>  select(target_lastauthor_year, type, raw_stat)

```

# Draft plot of effect sizes

* We will compare the original, 1st replication, and re-replication effect sizes, as well as any effect sizes coming from independent replications (where effect size can be computed)

```{r, fig.height=6, fig.width=8}
#test on just one
colors <- c("original"="black", "rep1"="#377EB8", "rescue"="#E41A1C", "additional"="#984EA3")
for_plotting <- parsed_d |> filter(is.na(calc_d_calc)&!is.na(calc_ES)) |> 
  mutate(point=calc_ES, low=calc_ES-1.96*calc_SE, high=calc_ES+1.96*calc_SE)

for_plotting_d <- parsed_d |> filter(!is.na(calc_d_calc)) |> 
  mutate(point=calc_d_calc, low=calc_d_calc-1.96*calc_d_calc_se, high=calc_d_calc+1.96*calc_d_calc_se)

orig_scale <- ggplot(for_plotting,aes(x=target_lastauthor_year,y=point,ymin=low,ymax=high, color=type, shape=type, group=desc(type)))+
  geom_errorbar(size=.5,width=.25, position=position_dodge(width=.4))+
  geom_point(position=position_dodge(width=.4))+
  coord_flip(ylim=c(-1.1,2))+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  scale_color_manual(values=colors)+
  labs(y="effect size on original scale", x="")

smd_scale <- ggplot(for_plotting_d,aes(x=target_lastauthor_year,y=point,ymin=low,ymax=high,shape=type, color=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
  coord_flip(ylim=c(-1.1,2))+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors)+
  theme(legend.position = "bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="")

plot_grid(orig_scale, smd_scale, nrow=2, rel_heights=c(.3,1))

```
# Subjective
* We will report the distribution of subjective replication success in our rescue sample

```{r}
subj <- parsed_d |> filter(type=="rescue") |> mutate(binary_success=ifelse(replication_score>.5,1,0))

subj |> group_by(replication_score) |> tally()

success <- filter(subj, binary_success==1)

```
* We will report the distribution of subjective replication success in our rescue sample

Of a total of `r nrow(subj)` replication, `r nrow(success)` succeeding at mostly or fully replicating the original results.

## We correlate all the things with subjective replication success

* The predictors used in https://osf.io/preprints/psyarxiv/dpyn6/ for exploratory analyses (the ones that need to be calculated)

```{r}
original <- parsed_d |> filter(type=="original") |> 
  rename_with(~str_c("original_",.), .cols=-target_lastauthor_year)

rep1 <- parsed_d |> filter(type=="rep1") |> select(target_lastauthor_year, rep_N=N)

for_cor <- subj |> left_join(original) |> left_join(rep1) |> 
  mutate(
  social=ifelse(subfield=="social", 1,0),
  other_psych=ifelse(subfield=="other-psych",1,0), 
  is_within=ifelse(within_between=="within", 1,0), 
  change_platform=ifelse(on_turk==original_on_turk, 0,1),
  log_trials=log(repeated_measure),
  log_sample=log(N),
  log_ratio_ss=log(N/original_N),
  rep_1_log_sample=log(rep_N),
  log_ratio_rep1_orig=log(rep_N/original_N),
  log_ratio_rescue_rep1=log(N/rep_N),
  open_data=ifelse(open_data=="yes",1,0),
  open_mat=ifelse(open_materials=="yes", 1,0),
  stanford=ifelse(stanford_internal=="yes",1,0)) |> 
  filter(!is.na(replication_score))

sub_cor <- function(var, stat = "estimate") {
  if (stat == "estimate") {
    cor.test(for_cor$replication_score, pull(for_cor, {{var}}))$estimate
  } else {
    cor.test(for_cor$replication_score, pull(for_cor, {{var}}))$p.value
  }
}

preds <- c("open_data",  "open_mat", "stanford", "change_platform", 
           "log_ratio_ss", "is_within", "single_vignette", "log_sample", 
           "log_trials", "social", "other_psych", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1")

cors <- tibble(preds = preds) |>
  mutate(r = sapply(preds, function(x) sub_cor(x, stat = "estimate")),
         p = sapply(preds, function(x) sub_cor(x, stat = "p"))) |>

  mutate(Predictors=factor(preds, levels=c("social", "other_psych", "is_within", "single_vignette", "change_platform", "open_data", "open_mat", "stanford", "pub_year", "log_trials", "log_sample", "log_ratio_ss", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1"), labels=c("Social", "Other psych", "Within subjects", "Single vignette", "Switch to online", "Open data", "Open materials", "Stanford", "Publication year", "Log trials", "Log original sample size", "Log rep/orig sample", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1"))) |> arrange(Predictors) |> select(Predictors, r, p)

library(kableExtra)
knitr::kable(cors, digits = 3, align='rcc') |> kable_styling(full_width=F, htmltable_class = "lightable-classic-2", font_size=40) 
#cors
```
* If there is a mixture of projects that succeed and fail to replicate the original results, we will qualitatively describe differences that may have played a role.

It looks like the ones with poor replication sample (due to inflated effect size, or exclusion/attrition/etc issues) where the rescue recruited more is the only sorta strong predictor

# table of expts by sample sizes & closenesses

```{r}
parsed_d |> select(target_lastauthor_year, type, N, closeness, replication_score) |> 
  filter(type!="additional") |> 
  pivot_wider(names_from="type", values_from=c(N, closeness, replication_score)) |> 
  select(paper=target_lastauthor_year, replication_score=replication_score_rescue, N_original, N_rep1, N_rescue,
         closeness_rep1, closeness_rescue) |> arrange(replication_score |> desc())

```
# PredInt and P_orig

```{r}
source(here("code","helper","stats.R"))
```

Compute prediction intervals and p_orig, write processed data to file.

```{r}
do_rma <- function(df) {
  if(!any(is.na(df$es))){
  r<- rma(yi=df$es, sei=df$es_se, slab=df$type)
  return(tibble(rma_est=r$beta[,1],rma_se=r$se))
  }
  return(tibble(rma_est=NA, rma_se=NA))
}

do_d_rma <- function(df) {
  if(!any(is.na(df$d))){
  r<- rma(yi=df$d, sei=df$d_se, slab=df$type)
  return(tibble(rma_d_est=r$beta[,1], rma_d_se=r$se))
  }
  return(tibble(rma_d_est=NA, rma_d_se=NA))

}

orig <- parsed_d |> filter(type=="original") |> select(target_lastauthor_year, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE)

do_p_orig <- function(es, es_se, rep_es, rep_se){
  if(!is.na(es)&!is.na(es_se)&!is.na(rep_es)&!is.na(rep_se)){
    return(Replicate::p_orig(es, es_se**2,rep_es, t2=0.21**2, rep_se**2))
  }
  return(NA)
}
foo <- parsed_d |> select(target_lastauthor_year, type, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE) |> 
  filter(type!="original") |> 
  group_by(target_lastauthor_year) |> 
  nest() |> 
  mutate(rma_result=map(data, do_rma),
         d_rma_result=map(data, do_d_rma)) |> 
  unnest_wider(rma_result) |> 
  unnest_wider(d_rma_result) |> 
  left_join(orig) |> 
  rowwise() |> 
  mutate(p_orig=do_p_orig(es, es_se, rma_est, rma_se),
         d_p_orig=do_p_orig(d, d_se, rma_d_est, rma_d_se))

```

* We will use p-original to evaluate how consistent the original effect size is with the totality of replications. We expect there to be a small number of replications, so we will impute the heterogeneity value as in https://osf.io/preprints/psyarxiv/dpyn6/.

## Secondary TODO you are here!

a) p-original between just the original and rescue,

b) p-original between the original and all replications except the rescue (in the case where no replications are found in the literature, this is the same as done in https://osf.io/preprints/psyarxiv/dpyn6/)

c) p-original between the rescue and all other replications.

We will visualize the consistency between original, 1st replication, rescue, and any other replications by plotting effect size and prediction interval for each.