### 2024-02-16 code/process.Rmd
finish first pass analysis

---
title: "251 Rescues"
output:
  html_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(viridis)
library(Replicate)
library(metafor)
library(esc)
library(here)
library(brms)
library(rstan)
library(googledrive)
library(glmnet)
library(tidybayes)
library(ggstance)
library("lattice")
library(reshape2)
library(ggrepel)
library(ggthemes)
library(knitr)
library(cowplot)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

theme_set(theme_bw())

model_location="code/models"
```

# Pull data

Download

```{r, eval=F}
f <- googledrive::as_dribble("https://docs.google.com/spreadsheets/d/12A4DblSbX_0tHP1mTVJhjNboI1YDngidkNWiSy0sN0g/edit#gid=0")
googledrive::drive_download(f, path=here("data","raw_data.xlsx"), overwrite=T)

raw_expts <- readxl::read_xlsx(here("data","raw_data.xlsx"), sheet="expt-level", skip=0)

raw_orig <- readxl::read_xlsx(here("data", "raw_data.xlsx"), sheet="original")

ready <- raw_expts |> left_join(raw_orig) |> write_csv(here("data", "combined_data.csv"))

```

Read in cache.

```{r}
d <- read_csv(here("data", "combined_data.csv")) |> 
  select(target_lastauthor_year, type, on_turk, repeated_measure, N, raw_stat, 
         same_direction, replication_score, closeness, subfield, target_year, stanford_internal, 
         open_data, open_materials, within_between, single_vignette)

```

# Parsing

We parse out values from the raw stats.

```{r}
source(here("code","helper","parse_stats.R"))
```

```{r}
parsed_d <- d |> 
  mutate(raw_stat=gsub(" ","",raw_stat),
         calc=pmap(list(raw_stat, within_between,N), do_parsing)) |> 
  unnest(cols=c(calc), names_sep="_") |> 
    mutate(
    calc_d_calc=case_when(
      type=="original" ~ abs(calc_d_calc),
      same_direction=="yes" ~ abs(calc_d_calc),
      same_direction=="no" ~ -abs(calc_d_calc),
      T ~ as.numeric(NA)
      ),
    calc_ES=case_when(
      type=="original" ~ abs(calc_ES),
      same_direction=="yes" ~ abs(calc_ES),
      same_direction=="no" ~ -abs(calc_ES),
      T ~ as.numeric(NA)
      ),
    type=factor(type, levels=c("original", "rep1", "rescue", "additional"))
    ) |> 
  rowwise()
```

## what didn't parse

Check that nothing that has a stat input and doesn't get an ES out.

```{r}

parsed_d |> filter(is.na(calc_ES)&(is.na(calc_d_calc))) |>  select(target_lastauthor_year, type, raw_stat)

parsed_d |> filter(is.na(calc_SE)&(is.na(calc_d_calc_se))) |>  select(target_lastauthor_year, type, raw_stat)

```

# Draft plot of effect sizes

* We will compare the original, 1st replication, and re-replication effect sizes, as well as any effect sizes coming from independent replications (where effect size can be computed)

```{r, fig.height=6, fig.width=8}
#test on just one
colors <- c("original"="black", "rep1"="#377EB8", "rescue"="#E41A1C", "additional"="#984EA3")
for_plotting <- parsed_d |> filter(is.na(calc_d_calc)&!is.na(calc_ES)) |> 
  mutate(point=calc_ES, low=calc_ES-1.96*calc_SE, high=calc_ES+1.96*calc_SE)

for_plotting_d <- parsed_d |> filter(!is.na(calc_d_calc)) |> 
  mutate(point=calc_d_calc, low=calc_d_calc-1.96*calc_d_calc_se, high=calc_d_calc+1.96*calc_d_calc_se)

orig_scale <- ggplot(for_plotting,aes(x=target_lastauthor_year,y=point,ymin=low,ymax=high, color=type, shape=type, group=desc(type)))+
  geom_errorbar(size=.5,width=.25, position=position_dodge(width=.4))+
  geom_point(position=position_dodge(width=.4))+
  coord_flip(ylim=c(-1.1,2))+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  scale_color_manual(values=colors)+
  labs(y="effect size on original scale", x="")

smd_scale <- ggplot(for_plotting_d,aes(x=target_lastauthor_year,y=point,ymin=low,ymax=high,shape=type, color=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
  coord_flip(ylim=c(-1.1,2))+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors)+
  theme(legend.position = "bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="")

plot_grid(orig_scale, smd_scale, nrow=2, rel_heights=c(.3,1))

```
# Subjective
* We will report the distribution of subjective replication success in our rescue sample

```{r}
subj <- parsed_d |> filter(type=="rescue") |> mutate(binary_success=ifelse(replication_score>.5,1,0))

tallied <- subj|> group_by(replication_score) |> tally()

tallied 
success <- filter(subj, binary_success==1)

d <- read_csv(here("data", "combined_data.csv")) |> filter(type=="rescue")

irr <- cor.test(d$TA_rep_score, d$MCF_rep_score, method="spearman")$estimate
irr
```
* We will report the distribution of subjective replication success in our rescue sample

Of a total of `r nrow(subj)` replication, `r nrow(success)` succeeding at mostly or fully replicating the original results (`r pluck(tallied,2,1)` with a score of 0, `r pluck(tallied, 2,2)` with a score of .75, and `r pluck(tallied, 2,3)` with a score of 1). The interrater reliability was `r irr |> round(2)`.

## We correlate all the things with subjective replication success

* The predictors used in https://osf.io/preprints/psyarxiv/dpyn6/ for exploratory analyses (the ones that need to be calculated)

```{r}
original <- parsed_d |> filter(type=="original") |> 
  rename_with(~str_c("original_",.), .cols=-target_lastauthor_year)

rep1 <- parsed_d |> filter(type=="rep1") |> select(target_lastauthor_year, rep_N=N)

for_cor <- subj |> left_join(original) |> left_join(rep1) |> 
  mutate(
  social=ifelse(subfield=="social", 1,0),
  other_psych=ifelse(subfield=="other-psych",1,0), 
  is_within=ifelse(within_between=="within", 1,0), 
  change_platform=ifelse(on_turk==original_on_turk, 0,1),
  log_trials=log(repeated_measure),
  log_sample=log(N),
  log_ratio_ss=log(N/original_N),
  rep_1_log_sample=log(rep_N),
  log_ratio_rep1_orig=log(rep_N/original_N),
  log_ratio_rescue_rep1=log(N/rep_N),
  open_data=ifelse(open_data=="yes",1,0),
  open_mat=ifelse(open_materials=="yes", 1,0),
  stanford=ifelse(stanford_internal=="yes",1,0)) |> 
  filter(!is.na(replication_score))

sub_cor <- function(var, stat = "estimate") {
  if (stat == "estimate") {
    cor.test(for_cor$replication_score, pull(for_cor, {{var}}))$estimate
  } else {
    cor.test(for_cor$replication_score, pull(for_cor, {{var}}))$p.value
  }
}

preds <- c("open_data",  "open_mat", "stanford", "change_platform", 
           "log_ratio_ss", "is_within", "single_vignette", "log_sample", 
           "log_trials", "social", "other_psych", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1")

cors <- tibble(preds = preds) |>
  mutate(r = sapply(preds, function(x) sub_cor(x, stat = "estimate")),
         p = sapply(preds, function(x) sub_cor(x, stat = "p"))) |>

  mutate(Predictors=factor(preds, levels=c("social", "other_psych", "is_within", "single_vignette", "change_platform", "open_data", "open_mat", "stanford", "pub_year", "log_trials", "log_sample", "log_ratio_ss", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1"), labels=c("Social", "Other psych", "Within subjects", "Single vignette", "Switch to online", "Open data", "Open materials", "Stanford", "Publication year", "Log trials", "Log original sample size", "Log rep/orig sample", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1"))) |> arrange(Predictors) |> select(Predictors, r, p)

library(kableExtra)
knitr::kable(cors, digits = 3, align='rcc') |> kable_styling(full_width=F, htmltable_class = "lightable-classic-2", font_size=40) 
#cors
```
* If there is a mixture of projects that succeed and fail to replicate the original results, we will qualitatively describe differences that may have played a role.

It looks like the ones with poor replication sample (due to inflated effect size, or exclusion/attrition/etc issues) where the rescue recruited more is the only sorta strong predictor

# table of expts by sample sizes & closenesses

```{r}
parsed_d |> select(target_lastauthor_year, type, N, closeness, replication_score) |> 
  filter(type!="additional") |> 
  pivot_wider(names_from="type", values_from=c(N, closeness, replication_score)) |> 
  select(paper=target_lastauthor_year, replication_score=replication_score_rescue, N_original, N_rep1, N_rescue,
         closeness_rep1, closeness_rescue) |> arrange(replication_score |> desc())

```
# PredInt and P_orig

```{r}
source(here("code","helper","stats.R"))
```

NOTE: can't do p-orig not on SMD scale here b/c tau imputation is in SMD units!!!

## orig v all on p-orig
```{r}
do_rma <- function(df) {
  if(!any(is.na(df$es))){
  r<- rma(yi=df$es, sei=df$es_se, slab=df$type)
  return(tibble(rma_est=r$beta[,1],rma_se=r$se))
  }
  return(tibble(rma_est=NA, rma_se=NA))
}

do_d_rma <- function(df) {
  if(!any(is.na(df$d))){
  r<- rma(yi=df$d, sei=df$d_se, slab=df$type)
  return(tibble(rma_d_est=r$beta[,1], rma_d_se=r$se))
  }
  return(tibble(rma_d_est=NA, rma_d_se=NA))

}

do_p_orig <- function(es, es_se, rep_es, rep_se){
  if(!is.na(es)&!is.na(es_se)&!is.na(rep_es)&!is.na(rep_se)){
    return(Replicate::p_orig(es, es_se**2,rep_es, t2=0.21**2, rep_se**2))
  }
  return(NA)
}

```

```{r}

orig <- parsed_d |> filter(type=="original") |> select(target_lastauthor_year, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE)

orig_v_others <- parsed_d |> select(target_lastauthor_year, type, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE) |> 
  filter(type!="original") |> 
  group_by(target_lastauthor_year) |> 
  nest() |> 
  mutate(rma_result=map(data, do_rma),
         d_rma_result=map(data, do_d_rma)) |> 
  unnest_wider(rma_result) |> 
  unnest_wider(d_rma_result) |> 
  left_join(orig) |> 
  rowwise() |> 
  mutate(d_p_orig=do_p_orig(d, d_se, rma_d_est, rma_d_se))

orig_v_others |> select(target_lastauthor_year, d_p_orig)

```

* We will use p-original to evaluate how consistent the original effect size is with the totality of replications. We expect there to be a small number of replications, so we will impute the heterogeneity value as in https://osf.io/preprints/psyarxiv/dpyn6/.

## Secondary TODO you are here!

## secondary, original and rescue
a) p-original between just the original and rescue,

```{r}
orig_v_rescue <- parsed_d |>   filter(type=="rescue") |> 
  select(target_lastauthor_year, type, res_d=calc_d_calc, res_d_se=calc_d_calc_se, res_es=calc_ES, res_es_se=calc_SE) |>

  left_join(orig) |> 
  rowwise() |> 
  mutate(d_p_orig=do_p_orig(d, d_se, res_d, res_d_se))

orig_v_rescue  |> select(target_lastauthor_year, d_p_orig)

```

## original v !rescue
b) p-original between the original and all replications except the rescue (in the case where no replications are found in the literature, this is the same as done in https://osf.io/preprints/psyarxiv/dpyn6/)

```{r}
orig_v_not_rescue <- parsed_d |> select(target_lastauthor_year, type, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE) |> 
  filter(type!="original", type!="rescue") |> 
  filter(target_lastauthor_year!="krauss2003") |> # have to avoid the divide by 0 error
  group_by(target_lastauthor_year) |> 
  nest() |> 
  mutate(rma_result=map(data, do_rma),
         d_rma_result=map(data, do_d_rma)) |> 
  unnest_wider(rma_result) |> 
  unnest_wider(d_rma_result) |> 
  left_join(orig) |> 
  rowwise() |> 
  mutate(d_p_orig=do_p_orig(d, d_se, rma_d_est, rma_d_se))

orig_v_not_rescue  |> select(target_lastauthor_year, d_p_orig)

```

## rescue v reps
c) p-original between the rescue and all other replications.

```{r}
rescue_only <- parsed_d |> filter(type=="rescue") |> select(target_lastauthor_year, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE)

rescue_v_reps <- parsed_d |> select(target_lastauthor_year, type, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE) |> 
  filter(type!="original", type!="rescue") |> 
  filter(target_lastauthor_year!="krauss2003") |> # have to avoid the divide by 0 error
  group_by(target_lastauthor_year) |> 
  nest() |> 
  mutate(rma_result=map(data, do_rma),
         d_rma_result=map(data, do_d_rma)) |> 
  unnest_wider(rma_result) |> 
  unnest_wider(d_rma_result) |> 
  left_join(rescue_only) |> 
  rowwise() |> 
  mutate(d_p_orig=do_p_orig(d, d_se, rma_d_est, rma_d_se))

rescue_v_reps |> select(target_lastauthor_year, d_p_orig)

```

## some predint viz! (this is currently with *no* hetereogeneity, idk what we actually want)
We will visualize the consistency between original, 1st replication, rescue, and any other replications by plotting effect size and prediction interval for each.

```{r}
do_predInt <- function(df){
  if(!is.na(df$es)&!is.na(df$es_se)&!is.na(df$rep_es)&!is.na(df$rep_es_se)){
    a <- Replicate::pred_int(df$es, df$es_se**2,df$rep_es, df$rep_es_se**2)
    return(tibble(low=a$int.lo, high=a$int.hi, inside=a$rep.inside))
  }
  return(tibble(low=NA,high=NA, inside=NA))
}

do_d_predInt <- function(df){
  if(!is.na(df$d)&!is.na(df$d_se)&!is.na(df$rep_d)&!is.na(df$rep_d_se)){
    a <- Replicate::pred_int(df$d, df$d_se**2,df$rep_d, df$rep_d_se**2)
    return(tibble(d_low=a$int.lo, d_high=a$int.hi, d_inside=a$rep.inside))
  }
  return(tibble(d_low=NA,d_high=NA, d_inside=NA))
}

```

```{r}
orig <- parsed_d |> filter(type=="original") |> select(target_lastauthor_year, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE)

predInts <- parsed_d |> 
  filter(type!="original") |> 
  group_by(target_lastauthor_year) |> 
  select(target_lastauthor_year, type, rep_d=calc_d_calc, rep_d_se=calc_d_calc_se, rep_es=calc_ES, rep_es_se=calc_SE) |> 
  left_join(orig) |> 
  group_by(target_lastauthor_year,type) |> 
  nest() |> 
  mutate(pred_int=map(data, do_predInt),
         d_pred_int=map(data,do_d_predInt)) |> 
  unnest_wider(pred_int) |> 
  unnest_wider(d_pred_int) |> 
  unnest_wider(data)

```

### viz attempts
```{r, fig.height=8, fig.width=6}
colors <- c("original"="black", "rep1"="#377EB8", "rescue"="#E41A1C", "additional"="#984EA3")

orig_scale <- ggplot(predInts |> filter(is.na(d_inside)&!is.na(inside)), aes(x=target_lastauthor_year, y=rep_es, ymin=low, ymax=high, color=type, group=desc(type), shape=type))+
  geom_errorbar(size=.5,width=.25, position=position_dodge(width=.4), color="black")+
  geom_point(position=position_dodge(width=.4))+
  geom_point(aes(y=es), color="black", position=position_dodge(width=.4), shape=1)+
  coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  scale_color_manual(values=colors)+
  labs(y="predInts on original scale", x="")

smd_scale <- ggplot(predInts |> filter(!is.na(d_inside)), aes(x=target_lastauthor_year, y=rep_d, ymin=d_low, ymax=d_high, color=type, group=desc(type), shape=type))+
  geom_errorbar(size=.5,width=.25, position=position_dodge(width=.4), color="black")+
  geom_point(position=position_dodge(width=.4))+
  geom_point(aes(y=d), color="black", position=position_dodge(width=.4), shape=1)+
  coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  scale_color_manual(values=colors)+
  labs(y="predInts on SMD scale", x="")

plot_grid(orig_scale, smd_scale, nrow=2, rel_heights=c(.3,1))

```
## predInt with heterogeneity

```{r}
do_d_predInt_het <- function(df){
  if(!is.na(df$d)&!is.na(df$d_se)&!is.na(df$rep_d)&!is.na(df$rep_d_se)){
    yio=df$d
    vio=df$d_se**2
    yir=df$rep_d
    vir=df$rep_d_se**2
    t2=.21**2
    pooled.SE = sqrt(vio + vir + t2)
    PILo.sens = yio - qnorm(0.975) * pooled.SE
    PIHi.sens = yio + qnorm(0.975) * pooled.SE
    PIinside.sens = (yir > PILo.sens) & (yir < PIHi.sens)
    return(tibble(d_low=PILo.sens, d_high=PIHi.sens, d_inside=PIinside.sens))
  }
  return(tibble(d_low=NA,d_high=NA, d_inside=NA))
}

```

```{r}
orig <- parsed_d |> filter(type=="original") |> select(target_lastauthor_year, d=calc_d_calc, d_se=calc_d_calc_se, es=calc_ES, es_se=calc_SE)

predInts_het <- parsed_d |> 
  filter(type!="original") |> 
  group_by(target_lastauthor_year) |> 
  select(target_lastauthor_year, type, rep_d=calc_d_calc, rep_d_se=calc_d_calc_se, rep_es=calc_ES, rep_es_se=calc_SE) |> 
  left_join(orig) |> 
  group_by(target_lastauthor_year,type) |> 
  nest() |> 
  mutate(d_pred_int=map(data,do_d_predInt_het)) |> 
  unnest_wider(d_pred_int) |> 
  unnest_wider(data)

```

### viz attempts
```{r, fig.height=8, fig.width=6}
colors <- c("original"="black", "rep1"="#377EB8", "rescue"="#E41A1C", "additional"="#984EA3")

 ggplot(predInts_het |> filter(!is.na(d_inside)), aes(x=target_lastauthor_year, y=rep_d, ymin=d_low, ymax=d_high, color=type, group=desc(type), shape=type))+
  geom_errorbar(size=.5,width=.25, position=position_dodge(width=.4), color="black")+
  geom_point(position=position_dodge(width=.4))+
  geom_point(aes(y=d), color="black", position=position_dodge(width=.4), shape=1)+
  coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  scale_color_manual(values=colors)+
  labs(y="predInts on SMD scale", x="")

```