### 2024-02-27 code/hopkins_es.Rmd
calculate raw mean sd for hopkins and reps

---
title: "Hopkins M & SD"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```
Key effect is that explantion level predicts rating (this is in an lm, but we're doing it outside for SMD here)

# Original

```{r}
dat <- read_csv(here("hopkins_raw.csv")) |>  # this has 147 unique IDs so seems likely to be the mturk sample post attention check?
  select(ID, rating, condition) |> 
  group_by(ID, condition) |> 
  summarize(rating=mean(rating)) |> 
  group_by(condition) |> 
  summarize(m=mean(rating), s=sd(rating))

  dat
```
# First rep
```{r}
## FROM https://github.com/psych251/hopkins2016/blob/master/writeup/hopkins2016_writeup.Rmd
raw_data <- read_csv("https://raw.githubusercontent.com/psych251/hopkins2016/master/writeup/data/hopkins2016_replication_data.csv")

#delete junk rows
raw_data <- raw_data[-c(1),]
raw_data <- raw_data[-c(1),]

#delete junk columns
raw_data <- raw_data %>%
  select(-(StartDate:Consent))
#add participant numbers
data_num <- raw_data %>%
  mutate(ID = 1:nrow(raw_data))

data_num <- data_num %>%
  select(-('Year':'Feedback'))
data_num <- data_num %>%
  select(Att_1, everything())

#convert to long format
data_long <- data_num %>%
  pivot_longer(cols = 'P1-HG_1':'S4-RB_1', names_to = "full_item", values_to = "rating", values_drop_na=TRUE)

#specify item attributes
data_long <- data_long %>%
  mutate(science = substr(full_item, start=1, stop=1))

data_long <- data_long %>%
  mutate(item = substr(full_item, start=1, stop=2))

data_long <- data_long %>%
  mutate(quality = substr(full_item, start=5, stop=5))

data_long <- data_long %>%
  select(-full_item)

data_long <- data_long %>%
  filter(rating!="")

#recode item names
data_long$quality <- ifelse(data_long$quality=="G", "good", "bad")

data_long$science <- ifelse(data_long$science=="P", "physics", data_long$science)
data_long$science <- ifelse(data_long$science=="C", "chemistry", data_long$science)
data_long$science <- ifelse(data_long$science=="B", "biology", data_long$science)
data_long$science <- ifelse(data_long$science=="N", "neuroscience", data_long$science)
data_long$science <- ifelse(data_long$science=="Y", "psychology", data_long$science)
data_long$science <- ifelse(data_long$science=="S", "social", data_long$science)

#convert columns to intended data types for analysis
data_long$rating <- as.integer(as.character(data_long$rating))

data_long$science <- factor(data_long$science, 
                  levels = c("physics", "chemistry", "biology", "neuroscience", "psychology", "social"),
                  labels = c("physics", "chemistry", "biology", "neuroscience", "psychology", "social"))
data_long$quality <- factor(data_long$quality, 
                  levels = c("good", "bad"),
                  labels = c("good", "bad"))
data_long$condition <- factor(data_long$Condition, 
                  levels = c("H", "R"),
                  labels = c("Horizontal", "Reductive"))

#exclude participants who failed attention check 1
data_excluded_both <- data_long %>%
  filter(Slider1_1==0)

#exclude participants who failed attention check 2
data_excluded_both <- data_excluded_both %>%
  filter(Att_1==3)
```
```{r}
data_excluded_both |>  select(ID, rating, condition) |> 
  group_by(ID, condition) |> 
  summarize(rating=mean(rating)) |> 
  group_by(condition) |> 
  summarize(m=mean(rating), s=sd(rating))
```
# Rescue

```{r}
raw_data <- read_csv("https://raw.githubusercontent.com/psych251/hopkins2016_rescue/main/data/replicationfull.csv")
## FROM https://github.com/psych251/hopkins2016/blob/master/writeup/hopkins2016_writeup.Rmd

#delete junk rows
raw_data <- raw_data[-c(1),]
raw_data <- raw_data[-c(1),]

#add participant numbers
data_num <- raw_data %>%
  mutate(ID = 1:nrow(raw_data))

data_num <- data_num %>%
  select(Att_1, everything())

#convert to long format
data_long <- data_num %>%
  pivot_longer(cols = 'P1-HG_1':'S4-RB_1', names_to = "full_item", values_to = "rating", values_drop_na=TRUE)

#specify item attributes
data_long <- data_long %>%
  mutate(science = substr(full_item, start=1, stop=1))

data_long <- data_long %>%
  mutate(item = substr(full_item, start=1, stop=2))

data_long <- data_long %>%
  mutate(quality = substr(full_item, start=5, stop=5))

data_long <- data_long %>%
  select(-full_item)

data_long <- data_long %>%
  filter(rating!="")

#recode item names
data_long$quality <- ifelse(data_long$quality=="G", "good", "bad")

data_long$science <- ifelse(data_long$science=="P", "physics", data_long$science)
data_long$science <- ifelse(data_long$science=="C", "chemistry", data_long$science)
data_long$science <- ifelse(data_long$science=="B", "biology", data_long$science)
data_long$science <- ifelse(data_long$science=="N", "neuroscience", data_long$science)
data_long$science <- ifelse(data_long$science=="Y", "psychology", data_long$science)
data_long$science <- ifelse(data_long$science=="S", "social", data_long$science)

#convert columns to intended data types for analysis
data_long$rating <- as.integer(as.character(data_long$rating))

data_long$science <- factor(data_long$science, 
                  levels = c("physics", "chemistry", "biology", "neuroscience", "psychology", "social"),
                  labels = c("physics", "chemistry", "biology", "neuroscience", "psychology", "social"))
data_long$quality <- factor(data_long$quality, 
                  levels = c("good", "bad"),
                  labels = c("good", "bad"))
data_long$condition <- factor(data_long$Condition, 
                  levels = c("H", "R"),
                  labels = c("Horizontal", "Reductive"))

#exclude participants who failed attention check 1
data_excluded_both <- data_long %>%
  filter(Slider1_1==0)

#exclude participants who failed attention check 2
data_excluded_both <- data_excluded_both %>%
  filter(Att_1==3)
```

```{r}
data_excluded_both |>  select(ID, rating, condition) |> 
  group_by(ID, condition) |> 
  summarize(rating=mean(rating)) |> 
  group_by(condition) |> 
  summarize(m=mean(rating), s=sd(rating))
```