### 2024-02-28 code/schechtman_es.Rmd
work on original schechtman ES

---
title: "Hopkins M & SD"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
library(R.matlab)
```
Key effect is that explantion level predicts rating (this is in an lm, but we're doing it outside for SMD here)

# Original

```{r}

dematlabify <- function(stuff) {
  foo <- stuff|> 
    as_tibble() |> 
    slice(2:n()) |> 
    mutate(across(everything(), ~map_vec(., ~ifelse(is.null(.), NA, .[[1]]))))

  names <- stuff[1,]|> map_vec(~.[[1]])
  colnames(foo) <- names

  return(foo)
}

# paper reports 39, but partway through reports "loss aversion" with 14, so I think maybe 25 in main analysis? That matches how many we have in thing1 at least

thing1 <- readMat(here("DataTest1.mat"))

ids <- dematlabify(thing1$BasicInfo)

dat1 <- dematlabify(thing1$DataStage1) # acquisition?

dat2 <- dematlabify(thing1$DataStage2) # testing

thing2 <- readMat(here("DataTest2.mat"))

ids2 <- dematlabify(thing2$BasicInfo) # has 6 22 13 2 11

dat3 <- dematlabify(thing2$DataSheet)

thing3 <- readMat(here("DataTest1_cases 4 13 18 omitted.mat")) # has 1-22

ids3 <- dematlabify(thing3$BasicInfo)

dat4 <- dematlabify(thing3$DataStage1) #acquisition

dat5 <- dematlabify(thing3$DataStage2) # testing

# not sure why 4 13 and 18 got excluded, but that seems to be what's reported in paper
```

```{r}
df <- dat5 |> left_join(ids3 |> select(IDNUM, `Tone Assignment nu.`)) |> 
  mutate(goodtone=ifelse(`Tone Assignment nu.`==1, 300,700),
         badtone=ifelse(`Tone Assignment nu.`==1, 700, 300)) |> 
  mutate(closest_distance=ifelse(abs(`absolute freq`-goodtone) > abs(`absolute freq`-badtone),abs(`absolute freq`-badtone), abs(`absolute freq`-goodtone) )) |> 
  filter(closest_distance<101) |> 
  mutate(closest_sound=ifelse(abs(`absolute freq`-goodtone) > abs(`absolute freq`-badtone),badtone, goodtone ),
         abs_diff=abs(closest_sound-`absolute freq`),
         generalized=ifelse(`Users choice: 1=positive 2=negative 3=other`==`closest sound: 1=good, 2=bad, 3=neither was closer than 200 Hz`,1,0),
         close_tone = ifelse(`closest sound: 1=good, 2=bad, 3=neither was closer than 200 Hz`==1, "good", "bad")) |> 
  select(IDNUM, goodtone, badtone, closest_sound, close_tone, abs_diff, generalized)

df |> nrow() #1584

# so 22 people, 3 rounds, 16 / round = 8 neg and 8 pos, if we exclude 0s
# or +4 in each category if we include 0s

#for rep 1420ish/20 = 71 so that's the same 72 if we include 0s

# for rescue

```

```{r}

mod <- lm(generalized ~ abs_diff*close_tone, df)

summary(mod)
```

```{r}

# figure out rescue
full_data <- read_csv(here("subject_data_final_full_processed.csv"))

# get all control questions
control_data = full_data %>%
  filter(trial_type == 'audio-button-response')

# check which participants did not answer "I can hear this" to 2 or more questions (button_pressed = 0)
exclude = control_data %>%
  filter(button_pressed != 0) %>%
  mutate(run_id = as.integer(run_id))

run_id_counts <- exclude %>%
  count(run_id)

run_id_counts_fin <- run_id_counts %>%
  filter(n != 1)

exclusion_participants = unique(run_id_counts_fin$run_id)

res <- full_data|> filter(run_id!=26) |> 
  filter(stage=="generalization") #|>
  #select(negative_tone, positive_tone, valence_match, distance, tone, key_responses, stimulus, run_id) |> 
  #filter(!is.na(distance))

nrow(res) #1512
```