### 2024-04-02 reconstruct_effect_sizes/denis_es.Rmd
add other stuff

---
title: "Denis"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```
We want an anova on general recognition of objects in wake v sleep x neg v neutral

It didn't come out so I checked other things

note that denis reports a hit - false alarm value but that very much isn't what payne does, so we try overall/general/specific on hit rate alone  to try to understand

```{r}
overall <- read_csv(here("reconstruct_effect_sizes/Expt1_OverallRecognition.csv")) |> 
  select(id, delay_cond, starts_with("hit")) |> 
  pivot_longer(starts_with("hit")) |> 
  rename(overall=value)

combined <- read_csv(here("reconstruct_effect_sizes/Expt1_SpecificRecognition.csv")) |> 
    select(id, delay_cond, starts_with("hit")) |> 
  pivot_longer(starts_with("hit")) |> 
  rename(specific=value) |> 
  left_join(overall) |> 
  separate(name, into=c(NA, "valence", "target")) |> 
  mutate(general=(overall-specific)/(1-specific))

test <- combined |> filter(!is.na(general))|> filter(! id %in% c("p1025", "p1032", "p1038", "p1050", "p1073", "p1104", "p1132", "p1239")) #can't handle NAs for general, so remove those participants

```

```{r}
# check 3-ways
library(ez)
anova_3way_obj = ezANOVA(
    #data=combined,
    data = test,
    #dv = .(overall),
    #dv = .(specific),
    dv= .(general),
    wid = .(id), 
    within = .(valence,target),
    between = .(delay_cond)
)

anova_3way_obj

#valence, target, valence x target on overall and specific
# cond x valence x target on specific only

#target, valence x target on general
```

```{r}

library(ez)
anova_2way_obj = ezANOVA(
    #data=combined |> filter(target=="object"),
    data = test |> filter(target=="object"),
    #dv = .(overall),
    #dv = .(specific),
    dv= .(general),
    wid = .(id), 
    within = .(valence),
    between = .(delay_cond)
)

anova_2way_obj

#valence on overall, specific, general

test |> filter(target=="object") |> group_by(valence, delay_cond) |> summarize(m=mean(general))
```