### 2024-08-09 manuscript/coverletter/collabra_response.tex
Mike's edits, some linking

\documentclass{stanfordletter}
%\makelabels

%\makelabels
\usepackage{todonotes}
\usepackage{varioref}
\usepackage{xr}
\externaldocument[paper-]{../revision}
\usepackage[american]{babel}
\usepackage{csquotes}
\usepackage[backend=biber,style=apa,doi=false,url=false,hyperref=true,apamaxprtauth=30,uniquename=false]{biblatex}
\usepackage{framed}
\DeclareLanguageMapping{american}{american-apa}
\newcommand{\citet}[1]{\textcite{#1}}
\newcommand{\citep}[1]{\parencite{#1}}

\usepackage{longtable,booktabs,array}
\usepackage{float}

\providecommand{\tightlist}{%
	\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\newcounter{section}

\newcommand{\section}[1]{
	\bigskip
	{\normalfont\Large\bfseries #1}}

\newcommand{\theysaid}[1]{\begin{leftbar} \noindent 
		\textsl{ #1}\end{leftbar}}

\newcommand{\revised}[1]{\begin{quote}	#1 \end{quote}}

\usepackage{hyperref}
\begin{document}
	\name{Veronica Boyce}
	\signature{\vspace{-35pt} Veronica Boyce, \\ on behalf of the author team}

	\begin{letter}{Professor Don van Ravenzwaaij \\ Editor-in-Chief \\ Collabra: Psychology }

		\opening{Dear Dr. van Ravenzwaaij,}

		\theysaid{I have now received two very thorough reviews of your manuscript, “Estimating the replicability of psychology experiments after an initial failure to replicate”, from experts in the field. I also independently read the manuscript before consulting these reviews. The reviewers (self-identified as Daniel Lakens and Rink Hoekstra) had mostly positive reactions to your manuscript. I agree that your manuscript has important strengths and also that there are some issues that need to be addressed. I therefore encourage you to submit a revised version for further consideration at Collabra: Psychology.}

		Thank you for your helpful comments! We have copied the reviews here and respond to them point by point, including quotes from the revised manuscript.

		We would like to reiterate our thanks both the editor and the reviewers for their thoughtful and prompt reviews. We hope that our changes to the manuscript and the documentation of our repository offer the necessary clarifications and nuance.

		\closing{Sincerely,}

	\section{Editor}

	\theysaid{To summarize what I perceive to be the main points of attention:}

	\theysaid{(1) data accessibility/availability: I shared reviewer \#1’s experience of several broken links, or links that had restricted access. I understand how this might happen with so many different projects, but if this article is to be published, all of these issues will need to be fixed. I strongly urge you, or one of your co-authors, to spend an afternoon clicking on all the links (and also on the links in the links: the preregistration documents include links that do not always work as well) and repair everything that does not work. I’m partial to reviewer \#1’s suggestion of a master-R-markdown, but I will not insist as long as everything works in the end.}

	%TODO check links

	We have checked that the links data/combined\_data.csv
	all work, and updated the README to be much more detailed.

	Unfortunately, we don't have a way to edit the individual pre-registrations, but we believe all the necessary information for the rescues is now available, linked from the data sheet.

	We also address reviewer 1's specific repository-related concerns more below; we suspect many of them would have been ameliorated had the reviewer been able to download the repository files (which, as noted below, are likely due to a known technical issue with github and Windows systems).

	\theysaid{(2) limitations/quality control: your discussion section, although admirably short, does not address any potential limitations your study has. I share reviewer \#1’s assessment that relative inexperience of the replicators is at least a potential limitation (but see https://journals.sagepub.com/doi/full/10.1177/1475725719857659, which you may wish to cite). There are several other points that reviewer \#1 is concerned about. You may not agree with all of them, so feel free to argue your case when you don’t, but across the board I do think a bit more engagement with these points is essential.}

	%TODO discussion section

	We have added an explicit ``Limitations'' subsection to the discussion, which we discuss further below in our response to reviewer 1.

	Also, see below for our changes and rebuttals to reviewer 1's other points.

	\theysaid{(3) implications: reviewer \#2 wonders how your results inform decisions on which papers to ‘rescue’. The final paragraph of the Discussion of your current submission only scratched the surface here. What to do? How valuable are these rescue attempts? Should we do more of them, or should we focus on ‘first’ replications, or should we not do them at all?}

	We have expanded the last part of the discussion section to acknowledge the much larger picture of different circumstances that motivate replications and rescues, and how these might be influenced by the prior beliefs of the investigators. We discuss the changes more below in response to reviewer 2's points.

	\theysaid{In your resubmission, please include a document with a point-by-point response to both the points I list here and the reviewers’ comments, outlining each change made in your manuscript or providing a suitable rebuttal. Please ensure that your revised files adhere to our author guidelines, and that the files are fully copyedited/proofed prior to upload. Please also ensure that all necessary copyright permissions have been obtained. This may be the last opportunity for major editing, therefore please fully check your file prior to re-submission.}

	\section{Reviewer 1}

	% \theysaid{The authors report the result of replication studies performed as part of an educational project. The results vary – some are relatively similar, of clear null replications, but others show a surprising amount of variability, possibly due to time and resource constraints, or a lack of experience. The main research question is what the probability is that failed studies can be rescued. The authors suggest this probability is rather low, but I think the current study does not really allow such generalizable conclusions. As Many Labs 5, the set of studies is not representative of the population. I have a number of observations that will hopefully help to improve the manuscript.}

	% \theysaid{Data needs to be shared in a reproducible manner}
	\theysaid{Regrettably I could not download the files from github – some filenames that have been uploaded are too long, and both cloning the repo as downloading at a zip file give errors (on my windows machine). }

	We are sorry this error occurred, and that reviewer 1 didn't realize the files could be downloaded directly from the OSF copy (linked in the paper).  The inability of reviewer 1 to download and open the files seems to have resulted in misunderstandings about the usability of files in the repo, which are much more understandable when opened on a computer.

	From googling, we believe the long files names issue to be a Windows issue, and to prevent this problem for future users we have added a note at the top of the README directing readers for how to address or sidestep the problem, namely:

	\revised{Note: If you are a windows user and get an error about overly long file names, please see \href{https://stackoverflow.com/questions/22575662/filename-too-long-in-git-for-windows}{this stackoverflow post} or \href{https://github.com/desktop/desktop/issues/17882} {this github issue} for how to fix it. Alternatively, you can download a zip of the repo from \href{https://osf.io/cyk5w/files/github}{the osf copy} by clicking "Download this folder".}

	Hopefully this warning will prevent future difficulties with accessing the repository.

	\theysaid{The repository is also not organized well enough to make it easy to see all the information. It needs to be restructured, with an actual read me, a clear folder structure, all files should be interoperable, etc. 
	}

	We think the folder structure is actually pretty clear (e.g., \texttt{first-replication-projects} directory contains the first replication projects), but we apologize that the README was not more detailed. We have rewritten the README to be much more explicit about how to reproduce the paper as well as what all the folders and files contain.

	See  \url{https://github.com/langcog/251-rescue-projects/tree/main} for the revised README.

	\theysaid{The reports are written in some sort of Rpubs format which consists of many files, but nothing I clicked lead to a readable file – I am quite certain this is not how these files should be exported and shared.}

	As we now clarify in the updated README, these are archival copies of the reports, which were originally posted to the Rpubs sharing website. From a cloned/downloaded copy of the repo, the html files can be opened in any web browser to see the reports.

	\theysaid{In the process.Rmd file, data is downloaded from a google drive https://docs.google.com/spreadsheets/d/12A4DblSbX\_0tHP1mTVJhjNboI1YDngidkNWiSy0sN0g/edit\#gid=0 which I have no access to. }

	It is common in RMarkdown files to add sections for documentation that do not need to be run every time to reproduce the result. Such sections are often marked \texttt{eval=F} and they do not evaluate. The piece of code referenced above is in such a section. \texttt{process.Rmd} both runs and knits without access to the file. However, to avoid future confusions, we have made the google spreadsheet openly viewable.

	\theysaid{It is clear this all needs to be checked by a reviewer before the manuscript can be accepted. I would strongly urge the author to add all files as reproducible markdown files (the manuscript, but also the individual data reports).}

	We apologize for any confusion about organization, but the manuscript is indeed written in an Rmd file that is fully  reproducible and is available in the repository. Each of the individual rescue projects was also written in a Rmd or Qmd reproducible report (included in the individual project repositories that are included in \texttt{individual-rescue-projects/}). We have now clarified this in the README.

	\theysaid{ The latter can also be exported as PDF is need by, with the raw data and code in the same folder. I believe papers like this require a high quality standard – I would want to see all details if this was a single study, and I want to see all detail of every study when this is a large scale replication project. Ideally, there is a single master file that 1) writes all reproducible original replication reports and rescue reports, 2) takes the output from these reports and compiles the manuscript that has been submitted.}

	We hope that our updated README now makes it clear where to find all of the details of all of the rescue projects as well as the sourcing for the overall manuscript.

	% \theysaid{Focus on the qualitative data}
	\theysaid{It is interesting to see the results of these projects, as they show the messiness of results when quality control is less strict than in other replication projects. In some ways, the current set of studies can be regarded as a lower bound of replication success under very challenging conditions. Moreover, the qualitative insights when describing the individual studies (which should be extended) is quite educational. For example, it shows the importance of sharing materials, and provides an impetus to require material sharing when a paper is published.}

	We agree that the frustration of original materials not being available or original instructions not being available is educational and hopefully serves as an impetus for more sharing.

	\theysaid{ I think a number of improvements to the manuscript are warranted, but I believe that the final result is worth sharing with the scientific community. Because there are no line numbers, the author will perhaps need to search a bit for where my comments apply.
		The case studies section is the most interesting part, but it should be extended to all studies.
		There are not too many studies, and some can be discussed briefly if results are consistent non-replications. This section should also proceed the quantitative analyses. It is so clear that some studies were incompetently performed that I personally no longer care about the quantitative analyses of these replications. Several of the studies with ugely diverging effect sizes are not discussed at all: Tarampi, Haimowitz, Porter (why did the rescue succeed?).}

	%TODO update after MS is potentially revised

	We thank the reviewer for this idea. We have added an Appendix with brief descriptions of the topics and outcomes of each of the studies.

	We disagree, however, about the ordering of quantitative and qualitative results. The quantitative results come directly from our preregistration and so should go first. In addition, we expect many readers will be more interested in the ``bottom line'' results than in the individual studies.

	% \theysaid{Be careful with generalizations}
	\theysaid{In the conclusions the authors generalize beyond the context of students doing replications, to investigators doing replications. However, I do not believe this project allows for such generalizations.}

	We appreciate the cautionary note, but we disagree about the generalizations that can be made from student work. In particular, the source of the disagreement may be the strong distinction the reviewer makes between between students and investigators. Perhaps this distinction comes from a misconception of who we are talking about when we talk about ``students,'' perhaps due to geographic or national differences in research training.

	To clarify, the replication rescues we report were performed primarily by first-year PhD students at an elite US institution, with a few more senior PhD students as well. These students were supervised and their work was reviewed by senior PhD students with research methods expertise and by the faculty instructor of the course. This supervision arrangement intentionally mirrors how research is typically conducted, even down to the multiple presentations of their replication projects that students give to the class (mirroring review and discussion in lab meetings). We expect that the reviewer would agree that many publications are in fact produced in exactly this way -- by PhD students, under the supervision of professors.

	We checked the google scholar pages of the student rescuers who are authors on our paper; 7/17 of them (40\%) already had peer-reviewed journal publications during the course, several of which were first-authored. These students are -- or will soon be -- publishing in the literature. We agree that they may not have extensive expertise, but we think they -- and thus their rescue projects -- are reasonably representative of work by early career researchers with limited resources.

	Since ``student'' can bring to mind different populations for different people, we have clarified in our limitations section who the rescuers were and how we think this is likely to generalize to other early career researchers (i.e. many of the people conducting research) but may not generalize to experts.

	\theysaid{The replication studies were often not done very well, or with severe limitations (lack of original materials, or switching to an online sample) and when we discuss whether an investigator should further pursuit a research line after a failed replications, it seems fair to at least first assume they will perform a very close replication (even closer than many studies reported here).}

	We disagree with the reviewer's characterization of our replication studies. There were some limitations for the original replications, including time limits and online samples. We believe these limits are common when early career researchers build on existing work: They may be limited in what resources (time, money, access to populations) are available, and so what results are accessible to them is relevant.

	As for when original materials aren't available, we believe that this limitation characterizes a vast amount of the replication literature. Unfortunately, materials availability is a severe issue. The extent to which this limitation affected our work is exactly the same as the extent to which it affects other investigators in non-student samples. Thus, if anything we would expect that this characteristic \emph{increases} the generality of our findings.

	% \theysaid{Experimental design of replication studies}
	\theysaid{The authors state they attempted to adequately power studies, but they refer to trying to achieve 2.5 times the sample size in the original study (ignoring the sample size and effect size in the replication study). But there are no details on what the original effect sizes were and how much power the replication studies had. There is only a mention of sample sizes, not of effect sizes. We only learn much later that several of the first replication studies had substantial challenges in recruiting sufficient participants, and had smaller sample sizes than the originals.}

	The idea behind 2.5x sample size is that it gives the replication an 80\% chance of correctly rejecting the null hypothesis of an effect that the original was 33\% powered for (see \href{https://doi.org/10.1177/0956797614567341}{Simonsohn, 2015, Psych Science}). This standard has the convenience of being agnostic to the type of experiment and test being conducted (unlike calculating effect sizes or calculating power, which both can be difficult to compute for complex designs and for cases where original data are not available). Per the Simonsohn proposal, the 2.5x multiplier is a target for answering the question, ``is there an effect big enough that the original study could have reasonably detected it.''

	% \theysaid{Subjective replication probability is unclear}
	\theysaid{The subjective replication probability coded as 0, 0.25, 0.75 and 1 is extremely confusing. It should be dropped.}

	We apologize, but we are not sure what is confusing about a holistic, rating-based metric for replication success. Indeed, the Replication Project in Psychology (Open Science Collaboration, 2015) included in its outcomes a subjective replication score. The current coding scheme was used in the paper that the current paper builds on ( \href{https://royalsocietypublishing.org/doi/full/10.1098/rsos.231240}{Boyce et al 2023}) because it was the coding done at the time of (most of) the original replications.

	There are many statistical metrics for what counts as a successful replication, including methods such as confidence interval overlap or the methods that we use in our quantitative results (e.g. p-original). The key issue with these measures is that they depend on the selection of a single key statistic for compariosn. As discussed in \href{https://journals.sagepub.com/doi/full/10.1177/2515245917740427}{Hawkins et al 2018} and Boyce et al 2023, original papers often report a number of effects, and it is not always clear which one is the most key. Additionally, often the theoretical pattern of interest -- as indicated by the authors -- is dependent on multiple numeric effects (ex. two main effects and their interaction). Thus, although we also analyze a key effect as well, we think the holistic, subjective assessment of whether or not a study replicated is an important part of our analysis.

	Finally, analysis of subjective replication ratings is one of our pre-registered analyses and so we do not believe it is appropriate to omit it.

	\theysaid{ When I look at the studies the authors seem to have used this score as some sort of count of properties of the original and replication. For example, a replication study that yielded a significant result but was statistically smaller must have gotten a .5 score. }

	We don't think this is a fair characterization of our rating scheme (which we have used in prior work, as discussed above).

	As we wrote the in the methods section, ``Each project was rated on the basis of subjective replication success (on a 0--1 scale) by both MCF and one of VB and BP.
	We thought about replication in terms of general match in key study results (direction, magnitude, and significance), rather than focusing on any singular numeric result or significance cut-off.''

	\theysaid{
		It would be more useful if the authors just listed the raw properties (was the study significant, was it statistically different from the original, etc). For example, the authors discuss the Krauss replication as succesful, but in Figure 1 it looks like the 95\% CI overlaps with 0. So, it was not statistically significant? Why is it considered to have replicated?}

	Whether a replication is statistically significant in the same direction as the original is one metric that is sometimes used as a measure of replicability, but, as noted by \href{https://journals.sagepub.com/doi/10.1177/1745691616646366}{Patil et al 2016}, it isn't a very good one: one study may have an effect size very similar to another but one might be p < .05 and the other might be p > .05.

	As we discuss in the case study section, the Krauss paper reported correct justification from 2/67 in the control condition and 13/34 in the guided thinking condition.  The rescue had 1/40  correct justifications in the control condition and 6/35 in the guided thinking condition. This effect was significant, though smaller than the effect in the original.

	Regarding Figure 1, the effect sizes shown are standardized mean differences (SMD; Cohen's $d$). In order to show effects on the same graph, we had to coerce all the numbers to be SMDs, even if this conversion required approximations (as in the case of the Krauss paper, where the results are proportions). For assessing the statistical significance of individual studies it makes more sense to look at their raw results, rather the CIs in Figure 1, which are approximately correct but not the best basis for inference.

	% \theysaid{Statistical analysis}
	\theysaid{Upon first read I was not too enthusiastic about using a replication metric that incorporated heterogeneity. It seems a weird choice to treat unknown variability across studies as statistical uncertainty. Surely, in the end a replication study should mean that we observe the same effect size, and that we understand the topic sufficiently to observe a similar effect repeatedly. However, Figure 1 convinced me that in this context something else needs to be done, as the variability is simple huge. In part this must be due to the students doing the replications. I wonder if the authors redid all analyses and checked for mistakes. Some differences are just too large. We also know from the published literature that there is fraud in student projects, and I wonder if this has been checked.}

	We note that the heterogeneity metric has been used frequently to quantify variation in replication success, notably in \href{doi.org/10.1073/pnas.2403490121}{Holzmeister et al. (2024)}, which provides a framework for thinking about heterogeneity as stemming from population variation, design variation, and analytic variation.

	Regarding other sources of heterogeneity, including error and fraud, we think these are relatively unlikely.  First, while we did not recode student analyses from scratch, teaching assistants reviewed student code several times, and remarked on stylistic issues and potential errors. Further, most analyses were quite straightforward.

	We are also quite confident that there is not fraud in our sample. Committing fraud would have been strictly more work than actually doing the work, and further, there was no incentive to do so. During the class, students submitted their projects for a number of checkpoints. Thus, we know that students had working implementations of the studies because the TAs went through every experiment and gave feedback. Additionally, students had to have working code that ran on their pilot data before they were approved for running the main data collection. We also know that the students ran studies on the class Prolific account because we can see how many participants were recruited and paid for each study. Thus, we know code was written, the experiment worked, and the data was collected. Students also knew that their grades in no way depended on what results they got (i.e. whether the study replicated), so had no incentive to change the data they had collected.

	Finally, don't you think that if students were going to commit fraud, they'd make their study work \emph{better}? Even the studies that did rescue had smaller effect sizes than the original studies.

	Finally, regarding student error: in our prior work, we have a larger sample of 176 student replication studies, and these studies found a similar rate of replication as previous studies such as RP:P (see Boyce et al 2023). If error rates were substantially different, we'd expect lower replication.

	\theysaid{For the additional replications it would be good to know if they were class projects or replications by more experienced researchers. For example, I assume at least one is from the RP:P. The authors could cite the studies that are in the literature, and indicate class projects with a different symbol in the figure than the triangle, or using an open triangle. }

	Thanks for this suggestion. In the case study section and the appendix, when we describe studies where there were additional replications, we now mention the additional studies and where they come from, citing those from the published literature.

	\theysaid{This needs a finer grained analysis – also because the authors choose to include these replications they are aware of in the meta-analysis to evaluate the results. If these replications are not registered reports, and they come from the published literature, their effects may be biased. If they are from the same authors, they might be biased. If they are done by students, they might be more variable due to some lack of expertise.}

	\theysaid{I do not really like the comparisons of the original against all replications, because sometimes there is so much variability between the studies, it feels weird to meta-analyze them, and it is an example of garbage in, garbage out.}

	We think there may be a misunderstanding here -- we aren't doing meta-analyses in order to aggregate effect size across the the original and replications.

	When we discuss the ``meta-analytic'' mean of some studies that is because the measures of consistency we use (p-original) is a measure of the consistency of two things. Thus, to compare more than two replications, some have to be pooled; we use meta-analysis to do the pooling to get an aggregate mean and standard error. As described in \href{https://rss.onlinelibrary.wiley.com/doi/full/10.1111/rssa.12572}{Mathur and VanderWeele 2020}, which introduces the p-original statistic, we uses the point estimate and standard error of the original compared to the average population effect size in the replications and standard error thereof. As they mention, the most common way to get the average population effect size is via a random effects meta analysis.

	\theysaid{First, I do not like the addition of additional replications. No systematic search has been done, so these studies might differ in some ways, and bias the overall effect size estimate. These additional studies can be mentioned, but, we should also see meta-analytic effect sizes without these replications.This can easily be presented in a table with different columns.}

	Students did search for direct replications of the studies, which if they are in the published literature are likely to be found. It is possible that there happen to be direct replications that aren't labeled as such that we could have missed, and of course, there could be unpublished or grey literature replications.

	We understand the concern about wanting to see just the replication and rescue, so we have added a column to Table 1 with this information.

	% New Table 1 shown below:

	% 	\begin{tabular}[t]{lrrrrr}
		% 		\toprule
		% 		\multicolumn{1}{c}{ } & \multicolumn{5}{c}{p-original comparing between} \\
		% 		\cmidrule(l{3pt}r{3pt}){2-6}
		% 		\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Original and} & \multicolumn{1}{c}{Rescue and} \\
		% 		\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-6}
		% 		Paper & All reps & 1st Rep and Rescue & Rescue & Non-rescue & Other reps\\
		% 		\midrule
		% 		Birch \& Bloom 2007 & 0.192 & 0.189 & 0.194 & 0.191 & 0.989\\
		% 		Child et al. 2018 & 0.669 & 0.669 & 0.641 & 0.858 & 0.778\\
		% 		Chou et al. 2016 & 0.055 & 0.055 & 0.101 & 0.005 & 0.231\\
		% 		Craig \& Richeson 2014 & 0.145 & 0.145 & 0.001 & 0.302 & 0.020\\
		% 		Gong et al. 2019 & 0.262 & 0.262 & 0.057 & 0.511 & 0.228\\
		% 		Haimovitz \& Dweck 2016 & 0.068 & 0.018 & 0.018 & 0.180 & 0.867\\
		% 		Hopkins et al. 2016 & 0.290 & 0.290 & 0.654 & 0.004 & 0.015\\
		% 		Jara-Ettinger et al. 2022 & 0.014 & 0.014 & 0.000 & 0.006 & 0.000\\
		% 		Krauss \& Wang 2003 & 0.271 & 0.271 & 0.519 & 0.139 & 0.311\\
		% 		Ngo et al. 2019 & 0.106 & 0.000 & 0.000 & 0.385 & 0.257\\
		% 		Paxton et al. 2012 & 0.011 & 0.011 & 0.005 & 0.023 & 0.649\\
		% 		Payne et al. 2008 & 0.023 & 0.071 & 0.031 & 0.079 & 0.895\\
		% 		Porter et al. 2016 & 0.370 & 0.370 & 0.905 & 0.002 & 0.003\\
		% 		Schechtman et al. 2010 & 0.712 & 0.712 & 0.654 & 0.770 & 0.877\\
		% 		Tarampi et al. 2016 & 0.000 & 0.000 & 0.000 & 0.000 & 0.145\\
		% 		Todd et al. 2016 & 0.062 & 0.100 & 0.124 & 0.058 & 0.778\\
		% 		Yeshurun \& Levy 2003 & 0.614 & 0.007 & 0.017 & 0.943 & 0.474\\
		% 		\bottomrule
		% 	\end{tabular}

	We again note that there appears to be a confusion over what is presented as we never present meta-analytic effect sizes for the reasons discussed above. We agree that such effect sizes wouldn't be easily interpretable.

	\theysaid{This can easily be presented in a table with different columns. Currently, only the p-values are reported – this is insufficient. We should see the meta-analytic effect sizes and their confidence intervals. I assume some are incredibly wide due to the widely varying effect sizes, and hence lead to a low p-value.}

	Again, it seems like there may be confusion here -- the ``p-values'' presented here are ``p-original.'' p-original is a p-value, in that it's a measure of how likely an estimate would be to be at least as extreme as observed given a ``null'' distribution. But it is a measure of statistical (in-)consistency of the effect sizes, not a measure of how close to zero the effect sizes are. It does not correspond to a classical meta-analytic hypothesis test.

	\theysaid{It would also be good to see measures of heterogeneity (even if tests for heterogeneity have low power with few studies). It is a bit weird all tests assume the same tau, when this is clearly not correct. I understand the idea – it is difficult to know the true tau – but this solution seems suboptimal.}

	We have too few datapoints to meaningfully estimate the heterogeneity from our data (Mathur and VanderWeele 2020 suggest using the calculated tau if there are 10 or more replications); however, we also know the heterogeneity is not 0, so our best option is to use an estimated measure of heterogeneity, as is done as a sensitivity analysis in \href{https://elifesciences.org/articles/71601}{Errington et al 2021} and Boyce et al 2023.

	% \theysaid{Minor points}
	\theysaid{Why is the median p-orig value reported? What does this tell us? I think this part of section 3.2 can be deleted.}

	P-original is a measure of the statistical (in-)consistency between the original and the replications. We present the median and IQR as a way to summarize the overall levels of inconsistency between original and replication in our sample of studies.

	\theysaid{Not all predictors are explained. What does the predictor ‘ Stanford’ mean? Or log trials? Please explain all predictors, otherwise it is no use to report them.}

	We apologize for this oversight. We have changed these row names to ``Original authors at Stanford'' and 
	``Log number of trials''. We have also amended the caption to clarify the reason for Stanford being relevant as well as provide a reference to where the predictors are discussed at length in Boyce et al 2023.

	% The revised caption reads:

	% \revised{Correlations between an individual predictor and the subjective replication score of the rescue project. The first set of predictors were pre-registered based on the correlates used in Boyce et al. (2023). Original authors at Stanford refers to at the time of the first replication as this may mean the first replication was not independent. For details about how the predictors were coded, see the Methods section of Boyce et al. (2023). The last three predictors were added post-hoc.}

	\theysaid{It is the open science collaborations, not consortium.}

	Oops, thank you for bringing this to our attention, we have fixed it.

	\theysaid{All studies replicated should be reported in the reference list.}
	We re-confirmed that all the original studies were present in the reference list.

	% \theysaid{Signed, Daniel Lakens}

	\section{Reviewer 2}

	% \theysaid{Thanks for the opportunity to review this paper.}

	% \theysaid{I believe this is a well-written paper, with an interesting question: Should investigators devote more time to continued replications?
		% In the paper, the authors try to “rescue” papers after an original failure to replicate. They show that in their relatively small sample of 17 studies, 5 could be rescued, and the others couldn’t, suggesting that rescuing studies has a high failure rate.}

	% \theysaid{I do believe this is an interesting finding, although not completely surprising: in a world in which it is already hard to replicate studies, it make sense that it is even harder to rescue an outcome after a failed replication attempt. But the effort of the authors shows this elegantly, and it also shows, maybe surprisingly to some, that studies actually can be rescued, despite failed replication attempts earlier. In the debates on these issues so far, a failed replication is sometimes presented as the end of the discussion, and indirectly this study shows that it doesn’t have to be. I think the data leave room for multiple sometimes seemingly contradictory conclusions: one could argue that we should be careful in spending our limited resources on trying to rescuing a study that was already unsuccessfully replicated, whereas another could say that the difficulty in finding the right subjects can sometimes affect results in such a way that we should be careful in taking the failed replication too seriously.}

	% \theysaid{So in general, I think the findings are interesting, and I think the authors did a good job in writing this up. What I am less convinced about is how this should inform decisions on rescuing papers. What does the general picture (a large majority of studies could not be rescued) tell me as a researcher who is contemplating rescuing a particular study? Since I assume that the selection of studies to be potentially rescued won’t be random, I probably already have reasons why I want to rescue this study. Either I am involved in the issue (I may be one of the people believing the original effect, or I may be even one of the authors of the original study), or I have reasons to believe that something went wrong in the replication study. Or I could believe that if the effect were true, this would be so important to know that it is worth taking the relatively large risk of a failure.}

	\theysaid{[...] the decision to re-replicate is most likely determined by many factors. The fact that on average (so, when picking a to be rescued study randomly) this probability is slim, should in my opinion only be a relatively small part of this decision. To me, this doesn’t disqualify this study, but I think it would be useful if the authors would acknowledge this somehow when discussing the implications of their findings.}

	This point is well taken and we have expanded our discussion to acknowledge the larger space of factors at play with replications and rescues and the motivations to do them.

	% \revised{Overall, the diminished and inconsistent effects we found in the rescues suggest that even if a re-replication ``works'', it may be difficult to build upon as follow-up studies will need large samples to detect small effects.
		% 	\newline

		% 	We have focused here on a narrow type of replication and rescue: that done by an early career researcher with the goal of establishing whether this is a result \emph{they} can build on. We acknowledge that this type of replication exists as part of a wider ecosystem of replications that have different properties and are done for different reasons.
		% 	\newline

		% 	In many cases, researchers may have strong prior beliefs about where a replication will succeed, and these priors may influence a researchers interpretation of a failed replication, and thus their calculations about where a re-replication will succeed or is worth attempting. For instance, a researcher who has high credence in the main result is more likely to attribute failure to specific details in the replication, and thus may want to do a rescue where those details are adjusted. On the other hand, a researcher who is dubious of the methods, results, or statistical approaches in an experiment, or who has previously failed to extend the experimental result, may be quicker to interpret a non-replication as a more definitive negative result. What a researcher chooses to do may also depend on the perceived strength of the replication, what size effects they think are relevant, and how important they think the target effect is. Thus, in some cases, when to re-replicate may be a question of prior beliefs and perceived utility rather than just likelihood of a re-replication succeeding.
		% 	\newline

		% 	We opened with a question about what an early-career psychology researcher should do given a failed replication: should they try again or give up and move on?
		% 	In our sample, the odds of a re-replication working are low (consistent with Ebersole et al., 2020).
		% 	Especially if there is not a clear, identifiable candidate cause for replication failure, it may be more efficient to select another result to build on.
		% }

	% \theysaid{One minor issue:}
	\theysaid{In the description of Figure 1, it is said that “the first replication’s key effect was non-zero”. I would expect all effects to be non-zero, so that’s not really informative. I guess what the authors mean is that supported for the original effect was found in the replication, or something along these lines}

	Thank you for this point, we have now clarified that what we meant was had a confidence interval not overlapping 0. 
	% The caption now reads

	% \revised{Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. In a few cases, the first replication's key effect was in the same direction as the original with a 95\% CI that did not overlap 0; however, in these cases the larger pattern of results was not consistent between original and first replication. }

	% \theysaid{So in general, I think this is a neat little paper, which, in my opinion, could benefit from adding a little bit more discussion/depth in the end on what its outcomes could entail.}

	% \theysaid{I always sign my reviews,
		% Rink Hoekstra}

		\end{letter}

\end{document}