### 2024-10-09 code/process.Rmd
trying again at link behavior

---
title: "251 Rescues"
output:
  html_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(viridis)
library(Replicate)
library(metafor)
library(esc)
library(here)
library(brms)
library(rstan)
library(googledrive)
library(glmnet)
library(tidybayes)
library(ggstance)
library("lattice")
library(reshape2)
library(ggrepel)
library(ggthemes)
library(knitr)
library(cowplot)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

theme_set(theme_bw())

model_location <- "code/models"
```

# Pull data

Download

```{r, eval=F}
f <- googledrive::as_dribble("https://docs.google.com/spreadsheets/d/12A4DblSbX_0tHP1mTVJhjNboI1YDngidkNWiSy0sN0g/edit#gid=0")
googledrive::drive_download(f, path = here("data", "raw_data.xlsx"), overwrite = T)

raw_expts <- readxl::read_xlsx(here("data", "raw_data.xlsx"), sheet = "expt-level", skip = 0)

raw_orig <- readxl::read_xlsx(here("data", "raw_data.xlsx"), sheet = "original")

ready <- raw_expts |>
  left_join(raw_orig) |>
  write_csv(here("data", "combined_data.csv"))

for_pretty <- ready |>
  knitr::kable(format = "pipe") |>
  write_lines(here("data", "combined_data.md"))
```

Read in cache.

```{r}
d <- read_csv(here("data", "combined_data.csv")) |>
  select(
    target_lastauthor_year, type, on_turk, repeated_measure, N, raw_stat,
    same_direction, replication_score, closeness, subfield, target_year, stanford_internal,
    open_data, open_materials, within_between, single_vignette, name_pretty
  )
```

# Parsing

We parse out values from the raw stats.

```{r}
source(here("code", "helper", "parse_stats.R"))
```

```{r}
parsed_d <- d |>
  mutate(
    raw_stat = gsub(" ", "", raw_stat),
    calc = pmap(list(raw_stat, within_between, N), do_parsing)
  ) |>
  unnest(cols = c(calc), names_sep = "_") |>
  mutate(
    calc_d_calc = case_when(
      type == "original" ~ abs(calc_d_calc),
      same_direction == "yes" ~ abs(calc_d_calc),
      same_direction == "no" ~ -abs(calc_d_calc),
      T ~ as.numeric(NA)
    ),
    calc_ES = case_when(
      type == "original" ~ abs(calc_ES),
      same_direction == "yes" ~ abs(calc_ES),
      same_direction == "no" ~ -abs(calc_ES),
      T ~ as.numeric(NA)
    ),
    type = factor(type, levels = c("original", "rep1", "rescue", "additional"))
  ) |>
  rowwise()
```

## what didn't parse

Check that nothing that has a stat input and doesn't get an ES out.

```{r}
parsed_d |>
  filter(is.na(calc_ES) & (is.na(calc_d_calc))) |>
  select(target_lastauthor_year, type, raw_stat)

parsed_d |>
  filter(is.na(calc_SE) & (is.na(calc_d_calc_se))) |>
  select(target_lastauthor_year, type, raw_stat)

# everything is parsing, yay!
parsed_d |> write_csv(here("data/parsed_data.csv"))
```

# Draft plot of effect sizes

* We will compare the original, 1st replication, and re-replication effect sizes, as well as any effect sizes coming from independent replications (where effect size can be computed)

```{r, fig.height=6, fig.width=8}
# test on just one
colors <- c("original" = "black", "rep1" = "#377EB8", "rescue" = "#E41A1C", "additional" = "#984EA3")
for_plotting <- parsed_d |>
  filter(is.na(calc_d_calc) & !is.na(calc_ES)) |>
  mutate(point = calc_ES, low = calc_ES - 1.96 * calc_SE, high = calc_ES + 1.96 * calc_SE)

for_plotting_d <- parsed_d |>
  filter(!is.na(calc_d_calc)) |>
  mutate(point = calc_d_calc, low = calc_d_calc - 1.96 * calc_d_calc_se, high = calc_d_calc + 1.96 * calc_d_calc_se)

orig_scale <- ggplot(for_plotting, aes(x = target_lastauthor_year, y = point, ymin = low, ymax = high, color = type, shape = type, group = desc(type))) +
  geom_errorbar(size = .5, width = .25, position = position_dodge(width = .4)) +
  geom_point(position = position_dodge(width = .4)) +
  coord_flip() +
  scale_size_area() +
  geom_hline(yintercept = 0, color = "black") +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) +
  scale_color_manual(values = colors) +
  labs(y = "effect size on original scale", x = "")

smd_scale <- ggplot(for_plotting_d, aes(x = target_lastauthor_year, y = point, ymin = low, ymax = high, shape = type, color = type, group = desc(type))) +
  geom_errorbar(size = .5, width = 0, position = position_dodge(width = .5)) +
  geom_point(position = position_dodge(width = .5)) +
  coord_flip(ylim = c(-1.1, 1.5)) +
  scale_size_area() +
  geom_hline(yintercept = 0, color = "black") +
  scale_color_manual(values = colors) +
  theme(
    legend.position = "bottom",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) +
  labs(y = "Standarized mean difference", x = "")

plot_grid(orig_scale, smd_scale, nrow = 2, rel_heights = c(.3, 1))
```

# Subjective

* We will report the distribution of subjective replication success in our rescue sample

```{r}
subj <- parsed_d |>
  filter(type == "rescue") |>
  mutate(binary_success = ifelse(replication_score > .5, 1, 0))

tallied <- subj |>
  group_by(replication_score) |>
  tally()

tallied
success <- filter(subj, binary_success == 1)

d <- read_csv(here("data", "combined_data.csv")) |> filter(type == "rescue")

irr <- cor.test(d$TA_rep_score, d$MCF_rep_score, method = "spearman")$estimate
irr
```

Of a total of `r nrow(subj)` replication, `r nrow(success)` succeeding at mostly or fully replicating the original results (`r pluck(tallied,2,1)` with a score of 0, `r pluck(tallied, 2,2)` with a score of .75, and `r pluck(tallied, 2,3)` with a score of 1). The interrater reliability was `r irr |> round(3)`.

## We correlate all the things with subjective replication success

* The predictors used in https://osf.io/preprints/psyarxiv/dpyn6/ for exploratory analyses (the ones that need to be calculated)

```{r}
original <- parsed_d |>
  filter(type == "original") |>
  rename_with(~ str_c("original_", .), .cols = -target_lastauthor_year)

rep1 <- parsed_d |>
  filter(type == "rep1") |>
  select(target_lastauthor_year, rep_N = N)

for_cor <- subj |>
  left_join(original) |>
  left_join(rep1) |>
  mutate(
    social = ifelse(subfield == "social", 1, 0),
    other_psych = ifelse(subfield == "other-psych", 1, 0),
    is_within = ifelse(within_between == "within", 1, 0),
    change_platform = ifelse(on_turk == original_on_turk, 0, 1),
    log_trials = log(repeated_measure),
    log_sample = log(N),
    log_ratio_ss = log(N / original_N),
    rep_1_log_sample = log(rep_N),
    log_ratio_rep1_orig = log(rep_N / original_N),
    log_ratio_rescue_rep1 = log(N / rep_N),
    open_data = ifelse(open_data == "yes", 1, 0),
    open_mat = ifelse(open_materials == "yes", 1, 0),
    stanford = ifelse(stanford_internal == "yes", 1, 0)
  ) |>
  filter(!is.na(replication_score))

sub_cor <- function(var, stat = "estimate") {
  if (stat == "estimate") {
    cor.test(for_cor$replication_score, pull(for_cor, {{ var }}))$estimate
  } else {
    cor.test(for_cor$replication_score, pull(for_cor, {{ var }}))$p.value
  }
}

preds <- c(
  "open_data", "open_mat", "stanford", "change_platform",
  "log_ratio_ss", "is_within", "single_vignette", "log_sample",
  "log_trials", "social", "other_psych", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1"
)

cors <- tibble(preds = preds) |>
  mutate(
    r = sapply(preds, function(x) sub_cor(x, stat = "estimate")),
    p = sapply(preds, function(x) sub_cor(x, stat = "p"))
  ) |>
  mutate(Predictors = factor(preds, levels = c("social", "other_psych", "is_within", "single_vignette", "change_platform", "open_data", "open_mat", "stanford", "pub_year", "log_trials", "log_sample", "log_ratio_ss", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1"), labels = c("Social", "Other psych", "Within subjects", "Single vignette", "Switch to online", "Open data", "Open materials", "Stanford", "Publication year", "Log trials", "Log original sample size", "Log rep/orig sample", "rep_1_log_sample", "log_ratio_rep1_orig", "log_ratio_rescue_rep1"))) |>
  arrange(Predictors) |>
  select(Predictors, r, p)

library(kableExtra)
knitr::kable(cors, digits = 3, align = "rcc") |> kable_styling(full_width = F, htmltable_class = "lightable-classic-2")
# cors
```
* If there is a mixture of projects that succeed and fail to replicate the original results, we will qualitatively describe differences that may have played a role.

It looks like the ones with poor replication sample (due to inflated effect size, or exclusion/attrition/etc issues) where the rescue recruited more is the only sorta strong predictor. (I added a couple non-pre-reg'd relative sample size of rep1 measures)

# table of expts by sample sizes & closenesses

```{r}
parsed_d |>
  select(target_lastauthor_year, type, N, closeness, replication_score) |>
  filter(type != "additional") |>
  pivot_wider(names_from = "type", values_from = c(N, closeness, replication_score)) |>
  select(
    paper = target_lastauthor_year, rescue_score = replication_score_rescue, N_original, N_rep1, N_rescue,
    closeness_rep1, closeness_rescue
  ) |>
  arrange(rescue_score |> desc())
```
# PredInt and P_orig

```{r}
source(here("code", "helper", "stats.R"))
```

NOTE: can't do p-orig not on SMD scale here b/c tau imputation is in SMD units!!!

## orig v all on p-orig
```{r}
do_rma <- function(df) {
  if (!any(is.na(df$es))) {
    r <- rma(yi = df$es, sei = df$es_se, slab = df$type)
    return(tibble(rma_est = r$beta[, 1], rma_se = r$se))
  }
  return(tibble(rma_est = NA, rma_se = NA))
}

do_d_rma <- function(df) {
  if (!any(is.na(df$d))) {
    r <- rma(yi = df$d, sei = df$d_se, slab = df$type)
    return(tibble(rma_d_est = r$beta[, 1], rma_d_se = r$se))
  }
  return(tibble(rma_d_est = NA, rma_d_se = NA))
}

do_p_orig <- function(es, es_se, rep_es, rep_se) {
  if (!is.na(es) & !is.na(es_se) & !is.na(rep_es) & !is.na(rep_se)) {
    return(Replicate::p_orig(es, es_se**2, rep_es, t2 = 0.21**2, rep_se**2))
  }
  return(NA)
}
```

```{r}
orig <- parsed_d |>
  filter(type == "original") |>
  select(target_lastauthor_year, d = calc_d_calc, d_se = calc_d_calc_se, es = calc_ES, es_se = calc_SE)

orig_v_others <- parsed_d |>
  select(target_lastauthor_year, type, d = calc_d_calc, d_se = calc_d_calc_se, es = calc_ES, es_se = calc_SE) |>
  filter(type != "original") |>
  group_by(target_lastauthor_year) |>
  nest() |>
  mutate(
    rma_result = map(data, do_rma),
    d_rma_result = map(data, do_d_rma)
  ) |>
  unnest_wider(rma_result) |>
  unnest_wider(d_rma_result) |>
  left_join(orig) |>
  rowwise() |>
  mutate(d_p_orig = do_p_orig(d, d_se, rma_d_est, rma_d_se))
```

* We will use p-original to evaluate how consistent the original effect size is with the totality of replications. We expect there to be a small number of replications, so we will impute the heterogeneity value as in https://osf.io/preprints/psyarxiv/dpyn6/.

## secondary, original and rescue
a) p-original between just the original and rescue,

```{r}
orig_v_rescue <- parsed_d |>
  filter(type == "rescue") |>
  select(target_lastauthor_year, type, res_d = calc_d_calc, res_d_se = calc_d_calc_se, res_es = calc_ES, res_es_se = calc_SE) |>
  left_join(orig) |>
  rowwise() |>
  mutate(d_p_orig = do_p_orig(d, d_se, res_d, res_d_se))
```

## original v !rescue
b) p-original between the original and all replications except the rescue (in the case where no replications are found in the literature, this is the same as done in https://osf.io/preprints/psyarxiv/dpyn6/)

```{r}
orig_v_not_rescue <- parsed_d |>
  select(target_lastauthor_year, type, d = calc_d_calc, d_se = calc_d_calc_se, es = calc_ES, es_se = calc_SE) |>
  filter(type != "original", type != "rescue") |>
  filter(target_lastauthor_year != "krauss2003") |> # have to avoid the divide by 0 error
  group_by(target_lastauthor_year) |>
  nest() |>
  mutate(
    rma_result = map(data, do_rma),
    d_rma_result = map(data, do_d_rma)
  ) |>
  unnest_wider(rma_result) |>
  unnest_wider(d_rma_result) |>
  left_join(orig) |>
  rowwise() |>
  mutate(d_p_orig = do_p_orig(d, d_se, rma_d_est, rma_d_se))
```

## rescue v reps
c) p-original between the rescue and all other replications.

```{r}
rescue_only <- parsed_d |>
  filter(type == "rescue") |>
  select(target_lastauthor_year, d = calc_d_calc, d_se = calc_d_calc_se, es = calc_ES, es_se = calc_SE)

rescue_v_reps <- parsed_d |>
  select(target_lastauthor_year, type, d = calc_d_calc, d_se = calc_d_calc_se, es = calc_ES, es_se = calc_SE) |>
  filter(type != "original", type != "rescue") |>
  filter(target_lastauthor_year != "krauss2003") |> # have to avoid the divide by 0 error
  group_by(target_lastauthor_year) |>
  nest() |>
  mutate(
    rma_result = map(data, do_rma),
    d_rma_result = map(data, do_d_rma)
  ) |>
  unnest_wider(rma_result) |>
  unnest_wider(d_rma_result) |>
  left_join(rescue_only) |>
  rowwise() |>
  mutate(d_p_orig = do_p_orig(d, d_se, rma_d_est, rma_d_se))
```

## show p_origs

```{r}
all <- orig_v_others |>
  select(target_lastauthor_year, orig_v_other = d_p_orig) |>
  left_join(orig_v_rescue |> select(target_lastauthor_year, orig_v_rescue = d_p_orig)) |>
  left_join(orig_v_not_rescue |> select(target_lastauthor_year, orig_v_not_rescue = d_p_orig)) |>
  left_join(rescue_v_reps |> select(target_lastauthor_year, rescue_v_reps = d_p_orig)) |>
  rename(paper = target_lastauthor_year)

all
```

## some predint viz! (this is currently with *no* hetereogeneity, idk what we actually want)
We will visualize the consistency between original, 1st replication, rescue, and any other replications by plotting effect size and prediction interval for each.

```{r}
do_predInt <- function(df) {
  if (!is.na(df$es) & !is.na(df$es_se) & !is.na(df$rep_es) & !is.na(df$rep_es_se)) {
    a <- Replicate::pred_int(df$es, df$es_se**2, df$rep_es, df$rep_es_se**2)
    return(tibble(low = a$int.lo, high = a$int.hi, inside = a$rep.inside))
  }
  return(tibble(low = NA, high = NA, inside = NA))
}

do_d_predInt <- function(df) {
  if (!is.na(df$d) & !is.na(df$d_se) & !is.na(df$rep_d) & !is.na(df$rep_d_se)) {
    a <- Replicate::pred_int(df$d, df$d_se**2, df$rep_d, df$rep_d_se**2)
    return(tibble(d_low = a$int.lo, d_high = a$int.hi, d_inside = a$rep.inside))
  }
  return(tibble(d_low = NA, d_high = NA, d_inside = NA))
}
```

```{r}
orig <- parsed_d |>
  filter(type == "original") |>
  select(target_lastauthor_year, d = calc_d_calc, d_se = calc_d_calc_se, es = calc_ES, es_se = calc_SE)

predInts <- parsed_d |>
  filter(type != "original") |>
  group_by(target_lastauthor_year) |>
  select(target_lastauthor_year, type, rep_d = calc_d_calc, rep_d_se = calc_d_calc_se, rep_es = calc_ES, rep_es_se = calc_SE) |>
  left_join(orig) |>
  group_by(target_lastauthor_year, type) |>
  nest() |>
  mutate(
    pred_int = map(data, do_predInt),
    d_pred_int = map(data, do_d_predInt)
  ) |>
  unnest_wider(pred_int) |>
  unnest_wider(d_pred_int) |>
  unnest_wider(data)
```

### viz attempts
```{r, fig.height=8, fig.width=6}
colors <- c("original" = "black", "rep1" = "#377EB8", "rescue" = "#E41A1C", "additional" = "#984EA3")

orig_scale <- ggplot(predInts |> filter(is.na(d_inside) & !is.na(inside)), aes(x = target_lastauthor_year, y = rep_es, ymin = low, ymax = high, color = type, group = desc(type), shape = type)) +
  geom_errorbar(size = .5, width = .25, position = position_dodge(width = .4), color = "black") +
  geom_point(position = position_dodge(width = .4)) +
  geom_point(aes(y = es), color = "black", position = position_dodge(width = .4), shape = 1) +
  coord_flip() +
  scale_size_area() +
  geom_hline(yintercept = 0, color = "black") +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) +
  scale_color_manual(values = colors) +
  labs(y = "predInts on original scale", x = "")

smd_scale <- ggplot(predInts |> filter(!is.na(d_inside)), aes(x = target_lastauthor_year, y = rep_d, ymin = d_low, ymax = d_high, color = type, group = desc(type), shape = type)) +
  geom_errorbar(size = .5, width = .25, position = position_dodge(width = .4), color = "black") +
  geom_point(position = position_dodge(width = .4)) +
  geom_point(aes(y = d), color = "black", position = position_dodge(width = .4), shape = 1) +
  coord_flip() +
  scale_size_area() +
  geom_hline(yintercept = 0, color = "black") +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) +
  scale_color_manual(values = colors) +
  labs(y = "predInts on SMD scale", x = "")

plot_grid(orig_scale, smd_scale, nrow = 2, rel_heights = c(.3, 1))
```

## predInt with heterogeneity

```{r}
do_d_predInt_het <- function(df) {
  if (!is.na(df$d) & !is.na(df$d_se) & !is.na(df$rep_d) & !is.na(df$rep_d_se)) {
    yio <- df$d
    vio <- df$d_se**2
    yir <- df$rep_d
    vir <- df$rep_d_se**2
    t2 <- .21**2
    pooled.SE <- sqrt(vio + vir + t2)
    PILo.sens <- yio - qnorm(0.975) * pooled.SE
    PIHi.sens <- yio + qnorm(0.975) * pooled.SE
    PIinside.sens <- (yir > PILo.sens) & (yir < PIHi.sens)
    return(tibble(d_low = PILo.sens, d_high = PIHi.sens, d_inside = PIinside.sens))
  }
  return(tibble(d_low = NA, d_high = NA, d_inside = NA))
}
```

```{r}
orig <- parsed_d |>
  filter(type == "original") |>
  select(target_lastauthor_year, d = calc_d_calc, d_se = calc_d_calc_se, es = calc_ES, es_se = calc_SE)

predInts_het <- parsed_d |>
  filter(type != "original") |>
  group_by(target_lastauthor_year) |>
  select(target_lastauthor_year, type, rep_d = calc_d_calc, rep_d_se = calc_d_calc_se, rep_es = calc_ES, rep_es_se = calc_SE) |>
  left_join(orig) |>
  group_by(target_lastauthor_year, type) |>
  nest() |>
  mutate(d_pred_int = map(data, do_d_predInt_het)) |>
  unnest_wider(d_pred_int) |>
  unnest_wider(data)
```

### viz attempts
```{r, fig.height=8, fig.width=6}
colors <- c("original" = "black", "rep1" = "#377EB8", "rescue" = "#E41A1C", "additional" = "#984EA3")

ggplot(predInts_het |> filter(!is.na(d_inside)), aes(x = target_lastauthor_year, y = rep_d, ymin = d_low, ymax = d_high, color = type, group = desc(type), shape = type)) +
  geom_errorbar(size = .5, width = .25, position = position_dodge(width = .4), color = "black") +
  geom_point(position = position_dodge(width = .4)) +
  geom_point(aes(y = d), color = "black", position = position_dodge(width = .4), shape = 1) +
  coord_flip() +
  scale_size_area() +
  geom_hline(yintercept = 0, color = "black") +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) +
  scale_color_manual(values = colors) +
  labs(y = "predInts on SMD scale", x = "")
```